---
title: "Movie Prediction" 
subtitle: "HU MATH 510 Final Presentation"
author: Jean, Gwenth, Joshua & Aaron
date: Apr 12, 2016
output: 
  ioslides_presentation: 
    highlight: haddock
    logo: photo.jpg
    smaller: yes
    widescreen: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
options(width = 1500)
```

```{r eval=TRUE, echo=FALSE}
require(dplyr)
require(ggplot2)
require(car)
require(tidyr)
require(magrittr)
require(lubridate)
require(stringr)
require(caret)
require(knitr)
```

```{r eval=TRUE, echo=FALSE}
load(url("https://stat.duke.edu/~mc301/data/movies.Rdata"))
```

## 1. Introduction

For this project, we are looking at *651* movies and generate features to predict the audience score showing on Rotten Tomato website. 

This presentation would walk through the exploratory and descriptive analysis of the data, process of feature engineering, model selection and finally evaluate the model on the movie **Spotlight**. 



## 2. EDA - 1

Plot the distribution of the audience_score for all the movies, and summarize. 

```{r echo=F, fig.height=3, fig.width=6}
qplot(audience_score, data=movies, geom = 'histogram')
```

```{r}
summary(movies$audience_score)
```

## 2. EDA - 2

What are the lowest rating and highest rating movies?

<br>

```{r}

movies %>%
  select(title, genre, runtime, studio, critics_rating, 
         critics_score, audience_rating, audience_score, imdb_rating) %>% 
  filter(audience_score %in% c(max(audience_score), min(audience_score))) %>% kable(format = "html")

```

<br>

Apprently *God Father Part II* is the highest rated movie. 

## 2. EDA - Univeriate or Bivariate Relationship 

```{r echo=F, fig.height=5, fig.width=6}

scatterplotMatrix(~audience_score+critics_score+runtime+imdb_num_votes+critics_rating, movies)

```


## 3. Inference

*Qestion: What's the difference between a Oscar nom movie and a non-nom one?*

```{r}

table(movies$best_pic_nom)

movies %>% 
  select(audience_score, critics_score, best_pic_nom) %>% 
  group_by(best_pic_nom) %>% 
  summarise(mean_aud_score = mean(audience_score), 
            means_crt_score = mean(critics_score)) %>% kable(format = "html")

```

## 3. Inference

*Qestion: Are the differences significant?*

```{r}
t.test(audience_score ~ best_pic_nom, alternative='less', data = movies) # yes, significantly smaller
t.test(critics_score ~ best_pic_nom, alternative='less', data = movies) # even more significantly different
```


## 4. Feature Engineering - 1

* In order to give the enough signal to predict the audience score, we generated many extra features other than initial ones:
* In the final inital model, we have the following engineered features:
+ `runtime_gt_90`: is runtime greater than 90?
+ `thtr_rel_wkd`: is the movie released in theater on Friday or Saturday? 
+ `thtr_rel_summer`: is the movie released in theater in summer (Jun, Jul)? 
+ `thtr_rel_holiday`: is the movie released in theater in holiday season (Oct, Nov, Dec)?
+ `thtr_rel_dump`: is the movie released in theater in Dump Months? (Jan, Feb, Aug, Sep)?
+ `thtr_dvd_rel_diff`: time difference between theater and dvd release. 
+ `dir_mv_freq`: how many movies did the director produced?
+ `std_mv_freq`: how many movies did the studio produced? 
+ `gr_mv_freq`: how many movies belong to this genre? 
+ ... continued in next slide

## 4. Feature Engineering - 2

Intuitively, one of the important feature of audience's perception of a video is:

* who starred in the movie? _(actors)_
* who directed the movie? _(director)_

So we engineered dummy variable for all the popular actors _(with more than 4 movies in the dataset)_ and popular dirctors _(with more than 2 movies directed in the dataset.)_

We use the following looped matrix to generate the dummy varaibles for popular actors and dirctors. 


```{r eval=F}
for(i in 1:651)  # for every movie (row)
{
  for(j in 1:37) # generate popular actor columns  
  {
    mymat[i,j] = as.numeric(str_detect(movie_model$actors[i], t[j]))
  }
}
```

We also interact `genre` with `dir_mv_freq`, `std_mv_freq` and `pop_act_dir` a measure of how many popular actors/dirctors this movie has. 


```{r eval=TRUE, echo=FALSE}
dir_mv_freq = as.data.frame(table(movies$director))
names(dir_mv_freq) = c("director", "dir_mv_freq")

std_mv_freq = as.data.frame(table(movies$studio))
names(std_mv_freq) = c("studio", "std_mv_freq")

gr_mv_freq = as.data.frame(table(movies$genre))
names(gr_mv_freq) = c("genre", "gr_mv_freq")

movies %>% 
  mutate(title_ws_ct = str_count(title, " ") + 1, # how many words in title?
         runtime_gt_90 = as.numeric(runtime>90), # is runtime greater than 90
         thtr_rel_date = ymd(str_c(thtr_rel_year, thtr_rel_month, thtr_rel_day, sep="-")), # date thrt release
         dvd_rel_date = ymd(str_c(dvd_rel_year, dvd_rel_month, dvd_rel_day,  sep="-")), # date dvd release 
         thtr_rel_wkd = as.numeric(wday(thtr_rel_date) %in% c(4,5)), # released in weekend?
         thtr_rel_summer = as.numeric(thtr_rel_month %in% c(6,7)), # released in summer? 
         thtr_rel_holiday = as.numeric(thtr_rel_month %in% c(10, 11, 12)), # released in holiday season?
         dvd_rel_holiday = as.numeric(dvd_rel_month %in% c(10, 11, 12)), # dvd released in holiday?
         thtr_rel_dump = as.numeric(thtr_rel_month %in% c(1, 2, 8, 9)), # released in 'dump month' (google it)
         thtr_dvd_rel_diff = as.numeric(dvd_rel_date-thtr_rel_date) # time diff btw dvd and thrt release
         ) %>% 
  unite(actors, actor1:actor5, sep=",") %>% # add actors concate column
  left_join(dir_mv_freq, by="director") %>% # add freq for dir
  left_join(std_mv_freq, by="studio") %>% # add freq for studio
  left_join(gr_mv_freq, by="genre") -> movie_model # add freq for gr. 

movie_model$actors %>% 
  str_c(collapse = ",") %>% 
  str_split(",") %>% 
  table(., exclude = "NA") %>% 
  as.data.frame() %>%
  arrange(desc(Freq)) -> act_mv_freq

movie_model$director %>% 
  table() %>% 
  as.data.frame() %>%
  arrange(desc(Freq)) -> dir_mv_freq_ls

act_mv_freq %>% 
  filter(Freq>4) %>% 
  select(.[1]) -> act_gt1_list

dir_mv_freq_ls %>% 
  filter(Freq>2) %>% 
  select(.[1]) -> dir_gt1_list

t = as.vector(act_gt1_list$.)
mymat = matrix(nrow=651, ncol=37)

for(i in 1:651)  # for each row
{
  for(j in 1:37) # for every actors with 
  {
    mymat[i,j] = as.numeric(str_detect(movie_model$actors[i], t[j]))
  }
}

t1 = paste0("dm_", tolower(gsub(" ", "_", t)))
t2 = str_replace_all(t1, fixed(".", T), "_")
t2 = str_replace_all(t2, fixed("'", T), "_")
t2 = str_replace_all(t2, fixed("-", T), "_")

##########################

d = as.vector(dir_gt1_list$.)
mymat2 = matrix(nrow=651, ncol=17)

for(i in 1:651)  # for each row
{
  for(j in 1:17) # for every actors with 
  {
    mymat2[i,j] = as.numeric(str_detect(movie_model$director[i], t[j]))
  }
}

d1 = paste0("dm_", tolower(gsub(" ", "_", d)))
d2 = str_replace_all(d1, fixed(".", T), "_")
d2 = str_replace_all(d2, fixed("'", T), "_")
d2 = str_replace_all(d2, fixed("-", T), "_")

dummy_dir = as.data.frame(mymat2)
names(dummy_dir) = d2

dummy_actors = as.data.frame(mymat)
names(dummy_actors) = t2

model_movie = cbind(movie_model, dummy_actors, dummy_dir)

model_movie$pop_act_dir = rowSums(model_movie[, c(42:95)], na.rm = T)
```

## 5. Model Fitting

Due to the bumber of predictors we have, we have to call our formula in a special way:

```{r eval=T, results="hide"}

base_formular = "audience_score ~ title_type + genre + runtime + mpaa_rating + thtr_rel_year
    + thtr_rel_month + imdb_num_votes + critics_rating + imdb_rating + genre*dir_mv_freq
    + genre*std_mv_freq + genre*gr_mv_freq + critics_score + best_pic_nom + best_pic_win + best_actor_win
    + best_actress_win + best_dir_win + top200_box + title_ws_ct + runtime_gt_90 + thtr_rel_wkd
    + thtr_rel_summer + thtr_rel_holiday + dvd_rel_holiday +  thtr_rel_dump + thtr_dvd_rel_diff + dir_mv_freq
    + std_mv_freq + gr_mv_freq + pop_act_dir + pop_act_dir*genre"

actor_formular = str_c(t2, collapse = " + ")
dir_formular = str_c(d2, collapse = " + ")

final_formular = as.formula(str_c(base_formular, actor_formular, dir_formular, sep = " + "))

fit = lm(final_formular, data = model_movie)

summary(fit)

```

## 5.Stepwise Selection to find the best model:

```{r eval=T, results="hide"}

fit2 = lm(final_formular, data = na.omit(model_movie))

step = step(fit2)

```

This is what the stepwise selection returns:

```{r eval=F}

lm(formula = audience_score ~ genre + thtr_rel_year + thtr_rel_month + 
    imdb_num_votes + critics_rating + imdb_rating + std_mv_freq + 
    best_pic_nom + runtime_gt_90 + thtr_dvd_rel_diff + pop_act_dir + 
    dm_john_travolta + dm_keanu_reeves + dm_val_kilmer + dm_charlize_theron + 
    dm_juliette_lewis + dm_naomi_watts + dm_robert_de_niro, data = na.omit(model_movie))

```

Notice that we didn't put `audience_rating` into the model because `audience_rating` is basically the categorical version of `audience_score`, which is kind of cheating... so we didn't use that. 

## 6. Final Model Fit

```{r eval=F}

fit3 = lm(audience_score ~ genre + thtr_rel_year + thtr_rel_month + 
    imdb_num_votes + critics_rating + imdb_rating + std_mv_freq + 
    best_pic_nom + runtime_gt_90 + thtr_dvd_rel_diff + pop_act_dir + 
    dm_john_travolta + dm_keanu_reeves + dm_val_kilmer + dm_charlize_theron + 
    dm_juliette_lewis + dm_naomi_watts + dm_robert_de_niro, data=model_movie)

```

So the final model returns the following significant predictors:

```{r eval=F}

genreArt House & International (++), genreDocumentary (+++), genreDrama(+), genreHorror(-), 
genreMusical & Performing Arts (+++), thtr_rel_year(---), imdb_num_votes(+++), critics_score(+++),
best_pic_nomyes(+), thtr_dvd_rel_diff(---), dm_nicolas_cage(-), dm_val_kilmer(+), dm_ben_affleck(+), dm_juliette_lewis(+)

```


Model has `Adjusted R-squared:  0.7829`
<br>
So apprentaly, genre is really significant, classic movies were rated higher and the more votes this movie get from imdb, the higer it is rated. 



